---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

## `bib2df` - Parse a BibTeX file to a tibble

Everyone writing reports and articles with LaTeX has probably used BibTeX before. BibTeX is the de facto standard for reference management and grounds its functionality on a list of references stored in local text file. Depending on the reference type, several fields are necessary to define a reference properly. An exemplary BibTeX entry looks as follows:

```
@Article{Binmore2008,
  Title = {Do Conventions Need to Be Common Knowledge?},
  Author = {Binmore, Ken},
  Journal = {Topoi},
  Year = {2008},
  Number = {1},
  Pages = {17--27},
  Volume = {27}
}
```

## Parse the BibTeX file to a tibble

The BibTeX format is not convenient for any kind of analysis or visualization. Many R applications require a `data.frame`, or its revised version, the `tibble`. `bib2df` offers a straightforward framework to parse a BibTeX file to a `tibble`.


```{r}
library(bib2df)
path <- system.file("extdata", "biblio.bib", package = "bib2df")
bib <- bib2df(path)
bib
```

## Installation

The latest version of `bib2df` can be installed from GitHub using `devtools::install_github()`:

```
devtools::install_github("ottlngr/bib2df")
```

A **CRAN** release is in the works.
